Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-29 sections/tensors.tex
okay?

\begin{itemize}
\item \textbf{Dense}
    The dense format is the simplest format, and simply maps $fiber(l, p)[i]
    \rightarrow fiber(l.lvl, p * l.shape + i)$. This format is used to store dense data,
    and is often a convenient format for the root level of a tensor. Because
    of its simplicity, freezing and thawing the level are no-ops.
\item \textbf{DenseRLE}
    The DenseRLE format is used to represent runs of repeated values.
    %
    Stores two vectors, $right$ and $ptr$, such that the $qth$ run in the $p^th$ subfiber begins and ends at $right[ptr[p] + q]$ and $right[ptr[p] + q + 1] - 1$, respectively.
    %
    Also stores a duplicate sublevel, $buf$, for deduplication.
\item \textbf{SparseList}
    The sparse list format is the simplest sparse format, and is used to construct the popular CSR, CSC, DCSR, DCSC, and CSF formats.
    %
    Stores two vectors, $idx$ and $ptr$, such that $idx[ptr[p] + q]$ is the index of the $qth$ nonzero in the subfiber at position $p$.
\item \textbf{SparsePinpoint}
    Because the SparsePinpoint format will only ever store one nonzero in each subfiber, we do not need the $ptr$ field.
    %
    Stores a vector $idx$ such that $idx[p]$ is the nonzero index in the subfiber at position $p$.
\item \textbf{SparseRLE}
    Similar to the DenseRLE level, but because the runs are stored on top of a background of fill values, we must also store the start of each run.
    %
    Stores three vectors, $left$, $right$, and $ptr$, such that the $qth$ run in the $p^th$ subfiber begins and ends at $left[ptr[p] + q]$ and $right[ptr[p] + q]$, respectively.
    %
    Also stores a duplicate sublevel, $buf$, for deduplication.
\item \textbf{SparseInterval}
    Similar to the SparseRLE level, but $ptr$ is redundant because we only store one run.
    %
    We also don't bother with deduplication here as we cannot store any intermediate results with duplicates.
    %
    Stores two vectors, $left$ and $right$, such that the run in subfiber $p$ begins and ends at $left[p]$ and $right[p]$ respectively.
\item \textbf{SparseVBL}
    The SparseVBL format is used to represent blocked data.
    %
    Stores three vectors, $idx$, $ptr$, and $ofs$, such that $ofs[ptr[p] + q]:ofs[ptr[p] + q + 1] - 1$ are the subpositions of block $q$ ending at index $idx[ptr[p] + q]$ in the subfiber at position $p$.
\item \textbf{SparseBand}
    Like SparseVBL, but only one block is stored in each subfiber.
    %
    Stores two vectors $idx$ and $ofs$, such that $ofs[p]:ofs[p + 1] - 1$ are the subpositions of the block ending at $idx[p]$ in subfiber $p$.
\end{itemize}

>>>>

\begin{table}[ht]
    \centering
    \begin{tabular}{p{14cm}}
    \hline
    \textbf{Sequential Reading and Updating} \\
    \hline
    \textbf{Dense}:
    The dense format is the simplest format, mapping $fiber(l, p)[i] \rightarrow fiber(l.lvl, p \times l.shape + i)$.
    %
    This format is used to store dense data and is often a convenient format for the root level of a tensor.
    %
    Due to its simplicity, freezing and thawing the level are no-ops. \\
    \textbf{DenseRLE}:
    Used to represent runs of repeated values, storing two vectors, $right$ and $ptr$, with $qth$ run in the $p^th$ subfiber starting and ending at $right[ptr[p] + q]$ and $right[ptr[p] + q + 1] - 1$, respectively.
    %
    It also stores a duplicate sublevel, $buf$, for deduplication. \\
    \textbf{SparseList}:
    The simplest sparse format, used to construct popular formats like CSR, CSC, DCSR, DCSC, and CSF.
    %
    It stores two vectors, $idx$ and $ptr$, such that $idx[ptr[p] + q]$ is the index of the $qth$ nonzero in the subfiber at position $p$. \\
    \textbf{SparsePinpoint}:
    Similar to SparseList, but only one nonzero in each subfiber, eliminating the need for the $ptr$ field.
    %
    It stores a vector $idx$, such that $idx[p]$ is the nonzero index in the subfiber at position $p$. \\
    \textbf{SparseRLE}:
    Similar to DenseRLE level, but because runs are sparse, we must also store the start of each run.
    %
    It stores three vectors $left$, $right$, and $ptr$, such that the $qth$ run in the $p^th$ subfiber begins and ends at $left[ptr[p] + q]$ and $right[ptr[p] + q]$, respectively.
    %
    It also stores a duplicate sublevel, $buf$, for deduplication. \\
    \textbf{SparseInterval}:
    Similar to SparseRLE, but only stores one run per subfiber, eliminating the need for the $ptr$ field.
    %
    This level does not deduplicate as it cannot store intermediate results with more than one run.
    %
    It stores two vectors, such that the run in subfiber $p$ begins and ends at $left[p]$ and $right[p]$ respectively. \\
    \textbf{SparseVBL}: Used to represent blocked data. It stores three vectors, $idx$, $ptr$, and $ofs$, such that $ofs[ptr[p] + q]:ofs[ptr[p] + q + 1] - 1$ are the subpositions of block $q$ ending at index $idx[ptr[p] + q]$ in the subfiber at position $p$. \\
    \textbf{SparseBand}: Similar to SparseVBL, but stores only one block per subfiber, eliminating the need for the $ptr$ field. It stores two vectors $idx$ and $ofs$, such that $ofs[p]:ofs[p + 1] - 1$ are the subpositions of the block ending at $idx[p]$ in subfiber $p$. \\
    \hline
    \textbf{Random Access Reading and Updating} \\
    \hline
    \textbf{SparseHash}:
    The sparse hash format uses a hash table to store the locations of nonzeros, and sorts the unique indices for iteration during the freeze phase.
    %
    This allows for efficient random access, but not incremental construction, as the freeze phase runs in time proportional to the number of nonzeros in the entire level.
    %
    It stores two vectors, $idx$ and $ptr$, such that $idx[ptr[p] + q]$ is the index of the $qth$ nonzero in the subfiber at position $p$. 
    %
    Also stores a hash table $tbl$ for construction and random access in the level. \\
    \textbf{SparseBytemap}
    The SparseBytemap format uses a bytemap to store which locations have been written to.
    %
    Unlike the SparseHash format, the bytemap assembles the entire space of possible subfibers.
    %
    This accelerates random access in the format, but requires a high memory overhead.
    %
    Because we don't want to reallocate all of the memory in each iteration, the declaration of this format instead re-assembles only the dirty locations in the tensor.
    %
    This format is analogous to the default workspace format used by TACO.
    %
    It stores two vectors, $idx$ and $ptr$, such that $idx[ptr[p] + q]$ is the index of the $qth$ nonzero in the subfiber at position $p$.
    %
    These vectors are used to collect dirty locations.
    %
    It also stores $tbl$, a dense array of booleans such that $tbl[shape * p + i]$ is true when there is a nonzero at index $i$ in the subfiber at position $p$. \\
    \hline
    \end{tabular}
    \caption{Description of Data Storage Formats}
    \label{table:formats}
\end{table}
